Just Add IFC! A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/retrofit.tex
rm useless stuff.
retrofit updates
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

Our overall plan will be to take a \textbf{{\color{red} target language}} with a formal
operational semantics and combine it with an \textit{{\color{blue} information
flow control language}}.  For concreteness, we use a simple, untyped,
ML-like language with mutable references as our target language.
However, our technique will be indifferent to the actual target language
chosen: it need only be prepared for interoperability by being expressed
in the style of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220},
and for simplicity, we require the semantics to be deterministic.
This procedure will
equip the target language with primitives for implementing dynamic
information control in the style of LIO~\cite{lio}, which we will
describe in more detail shortly.

@@

In Figure~\ref{fig:ml}, we give a simple, untyped ML-like language with
mutable references, prepared for combination with an
information flow control language.  The presentation is standard, but it
is worth paying some attention to the Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} used to define the
operational semantics of the system.  Our language defines an evaluation
context |tE|; however, the evaluation rules have been expressed in terms
of a different evaluation context |curlyRedE|,
following the approach of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220} to combine operational
semantics of multiple languages.  To
derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, bracket tE te|.
However, when we combine this language with an IFC language, we must
reinterpret the meaning of this evaluation context.
In general, we require that a target language be expressed in terms
of some global machine state |tS|, some evaluation context |tE|,
some expressions |te|, some set of values |tv| and a deterministic
reduction relation on full configurations $|tS| \times |tE| \times |te|$.

@@

\and
\inferrule[T-deref]
{ |memberf (ta, tv) tS| }
{| conf tS (!ta) -> conf tS (tv) |}

@@

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-threaded}
evaluation rules for a minimal information-flow control language.  The
evaluation context |curlyBlueE| represents the context
of the current task which is running, including on the right-hand side
the current label |il| (which protects the program counter of the task)
and the task ID |id|.

@@

%   \Red{When both languages have booleans, it is relatively simple to define
%   a natural embedding which preserves booleans across boundaries.}
%   \begin{mathpar}
%   \inferrule[I-bool]
%   { |b member {true, false}|  } {| iniE (IT b) -> iniE b |}
%   \and
%   \inferrule[T-bool]
%   { |b member {true, false}| } {| intE (TI b) -> intE b |}
%   \end{mathpar}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Our overall plan will be to take a \textbf{{\color{red} target language}} with a formal
operational semantics and combine it with an \textit{{\color{blue} information
flow control language}}.  While our technique will be indifferent to the
actual target language, for concreteness we give a simple, untyped,
ML-like language to illustrate some examples, as well as express
how the language should be prepared for interoperability
in the style of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220}.
We then equip this target language with primitives for implementing dynamic
information control in the style of LIO~\cite{lio}, which we will
describe in more detail shortly.

@@

In Figure~\ref{fig:ml}, we give a simple, untyped ML-like language with
mutable references, prepared for combination with an
information flow control language.  The presentation is standard, but it
is worth paying some attention to the Felleisen-Hieb reduction
semantics~\cite{Felleisen:1992:RRS:136293.136297} used to define the
operational semantics of the system.  Our language defines an evaluation
context |tE|; however, the evaluation rules have been expressed in terms
of a different evaluation context |curlyRedE|,
following the approach of Matthews and
Findler~\cite{Matthews:2007:OSM:1190216.1190220} to combine operational
semantics of multiple languages.  To
derive the usual operational semantics for this language, the evaluation
context merely needs to be defined as |conf tS te @= tS, bracket tE te|.
However, when we combine this language with an IFC language, we must
reinterpret the meaning of this evaluation context.
In general, we require that a target language be expressed in terms
of some global machine state |tS|, some evaluation context |tE|,
some expressions |te|, some set of values |tv| and a \emph{deterministic}
reduction relation on full configurations $|tS| \times |tE| \times |te|$.

@@

\and
\inferrule[T-deref]
{ |memberf ((ta, tv)) tS| }
{| conf tS (!ta) -> conf tS (tv) |}

@@

In Figure~\ref{fig:ifc}, we give the syntax and \emph{single-threaded}
evaluation rules for a minimal information-flow control language.  The
evaluation context |curlyBlueE| represents the context
of the current task which is running, which has three important pieces
of state: the global message queues |iS|,
the current label |il| (which protects the program counter of the task)
and the task ID |id|.

@@
